---
title: "Influence of the initialization"
author: Morgane Pierre-Jean
vignette: >
  %\VignetteIndexEntry{Influence of the initialization}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[UTF-8]{inputenc}
---
  
```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = TRUE)
```

In this document, we evaluate the performance of the various types of initialization.
  - Hclust
  - Random
  - SNF
  - SVD
  
To perform the evaluation, we simulate 10 datasets, then we compute the ARI and the percentage of the variance explained.


## Useful packages
  
```{r, library}
library(Matrix)
library(tidyverse)
library(PintMF)
library(CrIMMix)
library(future)
library(mclust)
```

## Simulate data with CrIMMix package

Four un balanced groups are simulated.
 
```{r, simulations}
nclust <- 4
nByclust= c(5, 10, 25, 10)
```


```{r, clust_func, message=FALSE}
clust_function <- function(R, nclust){
  clust <- R$W %>% dist %>% hclust(method="ward.D2") %>% cutree(nclust)
}
```

```{r, run_init, eval=FALSE}
set.seed(55)
perf <- do.call(rbind, lapply(1:10, function (ii){
  print(ii)
  c_1 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=100, prop=0.05, noise=0.2)
  c_2 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=50, flavor="binary",
                   params=list(c(p=0.6)), prop=0.1, noise=0.3)
  
  params_beta <- list(c(mean1=-1, mean2=1, sd1=0.5, sd2=0.5))
  
  c_3 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=500,flavor="beta", 
                   params=params_beta, prop=0.2, noise=0.2)
  data <- list(c_1$data, c_2$data, c_3$data)
  true.clust <- c_1$true.clusters
  data_names <- c("gaussian", "binary", "beta-like")
  names(data) <- data_names
  data_t <- data
  data_t[[3]] <- log2(data[[3]]/(1-data[[3]]))
  
  R_snf <- SolveInt(Y=data_t, p=4, max.it=5, verbose=FALSE, init_flavor="snf", flavor_mod="glmnet")
  
  R_svd <- SolveInt(Y=data_t, p=4, max.it=5, verbose=TRUE, init_flavor="svd", flavor_mod="glmnet")
  R_random <- SolveInt(Y=data_t, p=4, max.it=5, verbose=FALSE, init_flavor="random", flavor_mod="glmnet")
  R_hclust <- SolveInt(Y=data_t, p=4, max.it=5, verbose=FALSE, init_flavor="hclust", flavor_mod="glmnet")
  df_ari <- data.frame(sim=ii,ARI=c(adjustedRandIndex(clust_function(R_snf, nclust), true.clust), 
                                    adjustedRandIndex(clust_function(R_svd, nclust), true.clust),
                                    adjustedRandIndex(clust_function(R_random, nclust), true.clust),
                                    adjustedRandIndex(clust_function(R_hclust, nclust), true.clust)),
                       method=c("SNF", "SVD", "Random", "hclust"))
  
  df_pve <- data.frame( 
                       PVE=c(R_snf$pve, R_svd$pve, R_random$pve, R_hclust$pve), 
                       method= rep(c("SNF","SVD", "Random", "Hclust"),
                                   times=c(length(R_snf$pve),length(R_svd$pve), length(R_random$pve), length(R_hclust$pve) )))
  
  return(list(ari=df_ari, pve=df_pve))
}))
```



## Performance evaluation

```{r, violin_plot, eval=FALSE}
ARI_dat <-  do.call(rbind, perf[1:10])
g <- ARI_dat %>% ggplot(aes(x=method, fill=method, y=ARI))+geom_violin()+theme_bw()+theme(legend.position = "none", axis.text.x = element_text(size=15), axis.title.y = element_text(size=15), axis.text.y = element_text(size=10))+xlab("")+geom_point(alpha=0.5)
ggsave(filename = '../../Figs/eval_init.pdf', plot = g, width=7, height=5)
g
```


## Number of iterations

```{r, pve_plot, eval=FALSE}
pve_dat <- do.call(rbind, lapply(perf[11:20], function(g){
  g <- g %>% group_by(method) %>% mutate(it = row_number())
} ))

g_pve <- pve_dat%>% 
  ggplot(aes(x=it, y=PVE,fill=method)) +
  geom_smooth(se = TRUE, aes(colour=method))+theme_bw()+ theme(legend.position="right", axis.text.x = element_text(size=15),axis.title.y = element_text(size=15))+scale_x_continuous(
    breaks = 1:10)+xlab('iteration')
g_pve
ggsave(filename = '../../Figs/pve_init.pdf', plot = g_pve, width=7, height=5)

```



