---
title: "Review"
author: "Morgane Pierre-Jean"
date: "11/08/2021"
output: html_document
vignette: >
  %\VignetteIndexEntry{Review}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

To perform the evaluation, we simulate 20 datasets, then we compute the ARI and the percentage of the variance explained.


## Useful packages

```{r, library}
library(Matrix)
library(tidyverse)
library(PintMF)
library(CrIMMix)
library(future)
library(mclust)
```

## Simulate data with CrIMMix package

Four un balanced groups are simulated.

```{r, simulations}
nclust <- 4
nByclust= c(5, 10, 25, 10)
```


```{r, clust_func, message=FALSE}
clust_function <- function(R, nclust){
  clust <- R$W %>% dist %>% hclust(method="ward.D2") %>% cutree(nclust)
}
```

```{r, run_init, eval=TRUE}
set.seed(55)
perf <- do.call(rbind, lapply(1:20, function (ii){
  print(ii)
  c_1 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=1000, prop=0.02, noise=0.2)
  c_2 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=500, flavor="binary",
                   params=list(c(p=0.3)), prop=0.1, noise=0.3)
  
  params_beta <- list(c(mean1=-1, mean2=1, sd1=0.5, sd2=0.5))
  
  c_3 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=500,flavor="beta", 
                   params=params_beta, prop=0.2, noise=0.2)
  data <- list(c_1$data, c_2$data, c_3$data)
  true.clust <- c_1$true.clusters
  data_names <- c("gaussian", "binary", "beta-like")
  names(data) <- data_names
  data_t <- data
  data_t[[3]] <- log2(data[[3]]/(1-data[[3]]))
  
  R_snf <- SolveInt(Y=data_t, p=4, max.it=5, verbose=FALSE, init_flavor="snf", flavor_mod="glmnet")
  
  R_no_sparse <- SolveInt(Y=data_t, p=4, max.it=5, verbose=FALSE, init_flavor="snf", flavor_mod="no_sparse")
  
  df_ari <- data.frame(sim=ii,ARI=c(adjustedRandIndex(clust_function(R_snf, nclust), true.clust), 
                                    adjustedRandIndex(clust_function(R_no_sparse, nclust), true.clust)),
                       method=c("Sparse", "No_sparse"))
  
  df_pve <- data.frame( 
    PVE=c(R_snf$pve, R_no_sparse$pve), 
    method= rep(c("Sparse","No_sparse"),
                times=c(length(R_snf$pve),length(R_no_sparse$pve))))
  
  return(list(ari=df_ari, pve=df_pve))
}))
```



## Performance evaluation

```{r, violin_plot, eval=TRUE}
ARI_dat <-  do.call(rbind, perf[1:20])
g <- ARI_dat %>% ggplot(aes(x=method, fill=method, y=ARI))+geom_violin()+theme_bw()+theme(legend.position = "none", axis.text.x = element_text(size=15), axis.title.y = element_text(size=15), axis.text.y = element_text(size=10))+xlab("")+geom_point(alpha=0.5)
ARI_dat %>% group_by(method) %>% summarise(mean=mean(ARI))
```

No influence on the classification 


## Sparsity vs no sparsity 

```{r}
c_1 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=1000, prop=0.02, noise=0.2)
c_2 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=500, flavor="binary",
                 params=list(c(p=0.3)), prop=0.1, noise=0.3)

params_beta <- list(c(mean1=-1, mean2=1, sd1=0.5, sd2=0.5))

c_3 <- simulateY(nclust = nclust,  n_byClust = nByclust, J=500,flavor="beta", 
                 params=params_beta, prop=0.2, noise=0.2)
data <- list(c_1$data, c_2$data, c_3$data)
true.clust <- c_1$true.clusters
data_names <- c("gaussian", "binary", "beta-like")
names(data) <- data_names
data_t <- data
data_t[[3]] <- log2(data[[3]]/(1-data[[3]]))

R_snf <- SolveInt(Y=data_t, p=4, max.it=5, verbose=FALSE, init_flavor="snf", flavor_mod="glmnet")

R_no_sparse <- SolveInt(Y=data_t, p=4, max.it=5, verbose=FALSE, init_flavor="snf", flavor_mod="no_sparse")
```


Clearly the model with sparsity on H give similar results on classification but besides, selects  relevant variables.





```{r}
df_1 <- tibble(no_sparse_H = R_no_sparse$H[[1]] %>% as.numeric(),sparse_H = R_snf$H[[1]] %>% as.numeric() )
g1 <- df_1 %>% ggplot(aes(y = no_sparse_H, x = sparse_H))+geom_point (alpha= 0.5)+theme_bw()+ylab("no sparse  coefficients")+xlab("sparse coefficients")

df_2 <- tibble(no_sparse_H = R_no_sparse$H[[2]] %>% as.numeric(),sparse_H = R_snf$H[[2]] %>% as.numeric() )
g2 <- df_2 %>% ggplot(aes(y = no_sparse_H, x = sparse_H))+geom_point (alpha= 0.5)+theme_bw()+ylab("no sparse  coefficients")+xlab("sparse coefficients")

df_3 <- tibble(no_sparse_H = R_no_sparse$H[[3]] %>% as.numeric(),sparse_H = R_snf$H[[3]] %>% as.numeric() )
g3 <- df_3 %>% ggplot(aes(y = no_sparse_H, x = sparse_H))+geom_point (alpha= 0.5)+theme_bw()+ylab("no sparse  coefficients")+xlab("sparse coefficients")

gridExtra::grid.arrange(g1, g2, g3, ncol = 3)
```

